Box 1 - TDT Curves

### 2023-09-21 ms/ms.qmd
Better subtitles to grab reader

@@

The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates. While modelling approaches such as these are not a replacement for actual empirical data, they could be used to provide informed judgements on thermal tolerance in the absence of such information, but should be tested (using cross-validation) and refined when necessary.

Interestingly, there was a reasonably strong phylogenetic correlation between $CT_{max}$ and critical fertility limits in *Drosophilids* (r = `r posterior.mode(phylocor_ctmax_cfls)`, 95% CI: `r HPDinterval(phylocor_ctmax_cfls)[,"lower"]` to `r HPDinterval(phylocor_ctmax_cfls)[,"upper"]`), but not between $CT_{max}$, CFLs and the slope from TDT curves (cor($CT_{max}$, $z^{TDT}$) = `r posterior.mode(phylocor_ctmax_tdt)`, 95% CI: `r HPDinterval(phylocor_ctmax_tdt)[,"lower"]` to `r HPDinterval(phylocor_ctmax_tdt)[,"upper"]`; cor(CFLs, $z^{TDT}$) = `r posterior.mode(phylocor_tdt_cfls)`, 95% CI: `r HPDinterval(phylocor_tdt_cfls)[,"lower"]` to `r HPDinterval(phylocor_tdt_cfls)[,"upper"]`). These results suggest that sub-lethal measures, such as CFLs strongly coincide with a species' upper lethal limit. It would be interesting to understand if this is a general phenomena across more taxonomic groups.

>>>>

##### Evolutionary signatures of thermal tolerance in *Drosophilids*

@@

##### Evolutionary relationships can help understand thermal tolerance in species lacking thermal tolerance data
The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates. While modelling approaches such as these are not a replacement for actual empirical data, they could be used to provide informed judgements on thermal tolerance in the absence of such information, but should be tested (using cross-validation) and refined when necessary.

##### Static lethal limits are evolutionarily correlated with sub-lethal limits in *Drosophilids*
Interestingly, there was a reasonably strong phylogenetic correlation between $CT_{max}$ and critical fertility limits in *Drosophilids* (r = `r posterior.mode(phylocor_ctmax_cfls)`, 95% CI: `r HPDinterval(phylocor_ctmax_cfls)[,"lower"]` to `r HPDinterval(phylocor_ctmax_cfls)[,"upper"]`), but not between $CT_{max}$, CFLs and the slope from TDT curves (cor($CT_{max}$, $z^{TDT}$) = `r posterior.mode(phylocor_ctmax_tdt)`, 95% CI: `r HPDinterval(phylocor_ctmax_tdt)[,"lower"]` to `r HPDinterval(phylocor_ctmax_tdt)[,"upper"]`; cor(CFLs, $z^{TDT}$) = `r posterior.mode(phylocor_tdt_cfls)`, 95% CI: `r HPDinterval(phylocor_tdt_cfls)[,"lower"]` to `r HPDinterval(phylocor_tdt_cfls)[,"upper"]`). These results suggest that sub-lethal measures, such as CFLs strongly coincide with a species' upper lethal limit. It would be interesting to understand if this is a general phenomena across more taxonomic groups.
